---
title: "urbanaltlas_ULB"
author: "******** yuan"
date: "2019/11/11"
output: html_document
---

```{r setup, results='hide', message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("reshape2")
library("BBmisc")
library("h2o")
library("rgdal")
library("sp")
library("dplyr")
library("Metrics")
```

# to reproduce the methods provide in ULB paper, which use urban atlas as the lulc informaiton source and random forest to set the ratio of classes as the predictors.
```{r}
ulb_raw_features = read.csv("/home/z/********/london/ULB_ATLAS/ulb_raw_features.csv")
```

# for each moa object, complete the features by match the class name.
```{r}
moa_features_ULB = dcast(ulb_raw_features,msoa11cd~class_name,value.var = "area_ratio",sum)
# merge population to this varaibles.
moa_pop = get(load("/home/z/********/london/Rspace/pop/raw_pop_mid_london.Rdata"))
colnames(moa_pop)

moa_features_ULB_pop = merge(moa_features_ULB,moa_pop[c("AreaCodes","population")],by.x = "msoa11cd", by.y= "AreaCodes")
# normalize pop
moa_features_ULB_pop$population = normalize(moa_features_ULB_pop$population,method = "range",range = c(0, 1))

summary(moa_features_ULB_pop)
moa_data = subset(moa_features_ULB_pop,select = -c(msoa11cd))
colnames(moa_data)
```

# train on the moa level and get the hyper-parameters.

```{r}
#h2o.shutdown(prompt = TRUE)
h2o.no_progress()
h2o.init(max_mem_size = "20g")
# Create training (70%) and test (30%) sets for the AmesHousing::make_ames() data.
# Use set.seed for reproducibility
#set.seed(123)
moa_data.h2o = as.h2o(moa_data)
split <- h2o.splitFrame(moa_data.h2o,c(0.8))
train.h2o <- h2o.assign(split[[1]], "train.hex")   
test.h2o <- h2o.assign(split[[2]], "test.hex")  
y <- "population"
x <- setdiff(names(train.h2o), y)
```
# RF in moa level
# default setting base line model
```{r}
# training basic GBM model with defaults
h2o.fit1 <- h2o.randomForest(
  training_frame = train.h2o,        ## the H2O frame for training
  x = x,
  y = y,
  ntrees = 5000, 
  max_depth = 100,
  stopping_rounds = 10
)

# assess model results
h2o.fit1
```
# randomDiscrete search
```{r eval = FALSE}


# hyperparameter grid
hyper_grid.h2o <- list(
  ntrees      = seq(100, 8100, by = 1000),
  mtries      = c(-1,3,4),
  max_depth = c(10,50,100,200,400,600),
  sample_rate = c(.65,.70,0.75, .80),
  nbins       = seq(20, 50, by = 5)
)
# random grid search criteria
search_criteria <- list(
  strategy = "RandomDiscrete",
  stopping_metric = "rmse",
  stopping_tolerance = 0.005,
  stopping_rounds = 10,
  max_runtime_secs = 60*60
  )

# build grid search 
random_grid <- h2o.grid(
  algorithm = "randomForest",
  grid_id = "rf_grid2",
  x = x, 
  y = y, 
  training_frame = train.h2o,
  hyper_params = hyper_grid.h2o,
  search_criteria = search_criteria
  )

# collect the results and sort by our model performance metric of choice
grid_perf2 <- h2o.getGrid(
  grid_id = "rf_grid2", 
  sort_by = "rmse", 
  decreasing = FALSE
  )
print(grid_perf2)

```
# best model
```{r eval = FALSE}
# Grab the model_id for the top model, chosen by validation error
best_model_id <- grid_perf2@model_ids[[1]]
best_model <- h2o.getModel(best_model_id)

# Now let's get performance metrics on the best model
h2o.performance(model = best_model, valid = TRUE)
```



# retrain the best model

```{r}
# train final model
h2o.final <- h2o.randomForest(
  x = x,
  y = y,
  nfolds = 5,
  training_frame = train.h2o,
  mtries = 4,
  nbins = 30,
   ntrees = 4100, 
  max_depth = 10,
  sample_rate = 0.75,
  stopping_rounds = 20#,
  #seed = 123
  
)

# model stopped after xx trees
h2o.final@parameters$ntrees

# cross validated RMSE
h2o.rmse(h2o.final, xval = TRUE)
```

# varaibles important
```{r}
h2o.varimp_plot(h2o.final, num_of_features = 10)
```

# prediction in moa's test dataset.
```{r}
# evaluate performance on new data
h2o.performance(model = h2o.final, newdata = test.h2o)

test_moa_predict = as.data.frame(h2o.predict(h2o.final, test.h2o))

predict_pop = test_moa_predict$predict
real_pop =  as.data.frame(test.h2o)$population
# rmse
rmse(real_pop,predict_pop)

# %rmse
(rmse(real_pop,predict_pop)*length(real_pop))/sum(real_pop)

#mae
mae(real_pop,predict_pop)

# %mae
(mae(real_pop,predict_pop) * length(real_pop))/sum(real_pop)

```

# prepare the sbl features
```{r}
sbl_data_features= read.csv("/home/z/********/london/ULB_ATLAS/ulb_raw_features_sbl.csv")
colnames(sbl_data_features)


dim(sbl_data_features)
sbl_features_ULB = dcast(sbl_data_features,sbl_id~class_name,value.var = "area_ratio",sum,fill = 0)
dim(sbl_features_ULB)
summary(sbl_features_ULB)

sbl_data = subset(sbl_features_ULB,select = -c(sbl_id))
```




# test directly apply on sbl with scaled prediction
```{r}
# convert test set to h2o object
sb_transferred.h2o <- as.h2o(sbl_data)
sb_transferred_predict = as.data.frame(h2o.predict(h2o.final, sb_transferred.h2o))
summary(sb_transferred_predict$predict)

```



# directly implement the model trained on moa to predict on sbl 

#3. test different aggreagation methods 
apply the restriction of moa 

```{r}

sb_shape <- readOGR("/home/z/********/london/ULB_ATLAS/street_blocksl_withinloa_gid.shp")

sbl_predicted_data = data.frame(pred_sbl_pop=sb_transferred_predict$predict, sb_id = sbl_features_ULB$sbl_id)

sbl_predicted_data_df = merge(sbl_predicted_data,sb_shape@data[c("SBL_ID","LSOA11CD")],by.x = "sb_id", by.y= "SBL_ID", all.x = TRUE)
colnames(sbl_predicted_data_df)

london_low <- readOGR("/home/z/********/london/LSOA_2011_London_3857.shp")
london_low_df = london_low@data[,c(1,3)]

sbl_predicted_df_lm = merge(sbl_predicted_data_df,london_low_df, by= "LSOA11CD", all.x = TRUE)



# aggreagte sb to moa
sbl_predicted_df_lm_agg = sbl_predicted_df_lm %>%
  group_by(MSOA11CD) %>% 
  summarise(predicted_pop_sum_moa = sum(pred_sbl_pop))

raw_pop_moa= get(load("/home/z/********/london/Rspace/pop/raw_pop_mid_london.Rdata"))
raw_pop_moa = raw_pop_moa[,c(1,3)]
colnames(raw_pop_moa) = c("MSOA11CD","moa_true_population")
summary(raw_pop_moa)

sbl_predicted_df_lm_agg_temp = merge(sbl_predicted_df_lm_agg,raw_pop_moa,by = "MSOA11CD",all.x = TRUE)
sbl_predicted_df_lm_agg_temp["ad_ratio"] = sbl_predicted_df_lm_agg_temp$moa_true_population/sbl_predicted_df_lm_agg_temp$predicted_pop_sum_moa
# adjust by pop in moa
sbl_predicted_df_adj = NULL
sbl_predicted_df_adj = merge(sbl_predicted_df_lm,sbl_predicted_df_lm_agg_temp, by= "MSOA11CD", all.x = TRUE)

sbl_predicted_df_adj$pred_adj_pop = sbl_predicted_df_adj$pred_sbl_pop * sbl_predicted_df_adj$ad_ratio

temp_final = merge(sb_shape@data,sbl_predicted_df_adj[,c("pred_sbl_pop","pred_adj_pop","sb_id")],by.x = "SBL_ID",by.y="sb_id",all.x=TRUE)

sb_shape@data=temp_final
writeOGR(sb_shape,dsn ="/home/z/********/london/ULB_ATLAS",layer = "predicted_adj_sbl",driver = "ESRI Shapefile")
```
```{r}
summary(sbl_predicted_df_adj$pred_adj_pop)
plot(density(sbl_predicted_df_adj$pred_adj_pop))
```

# aggreagete to loa and output the aggregated value on loa and the spatial residual
```{r}
# aggreagete to loa
sbl_predicted_df_adj_loa = sbl_predicted_df_adj %>% 
  group_by(LSOA11CD) %>% 
  summarise(pred_loa_pop = sum(pred_adj_pop))

# merge with true pop in loa 
raw_pop_loa= get(load("/home/z/********/london/Rspace/pop/raw_pop_london_low.Rdata"))
raw_pop_loa["LSOA11CD"] = raw_pop_loa$AreaCodes
sbl_predicted_df_adj_loa = merge(sbl_predicted_df_adj_loa,raw_pop_loa, by= "LSOA11CD", all.x = TRUE)
colnames(sbl_predicted_df_adj_loa)
sbl_predicted_df_adj_loa = sbl_predicted_df_adj_loa[,c(1:5)]


colnames(sbl_predicted_df_adj_loa)[5] = "loa_true_pop"
sbl_predicted_df_adj_loa["residual"] = sbl_predicted_df_adj_loa$loa_true_pop-sbl_predicted_df_adj_loa$pred_loa_pop
summary(sbl_predicted_df_adj_loa)

london_low_shp = merge(sbl_predicted_df_adj_loa,london_low@data,by = "LSOA11CD",all.y= TRUE)
london_low@data = london_low_shp
writeOGR(london_low,dsn ="/home/z/********/london/ULB_ATLAS",layer = "predicted_adj_sbl_loa",driver = "ESRI Shapefile")
```

#3.  esitmate accuracy 
```{r}
predict_pop = sbl_predicted_df_adj_loa$pred_loa_pop
real_pop =  sbl_predicted_df_adj_loa$loa_true_pop

# rmse
rmse(real_pop,predict_pop)

# %rmse
(rmse(real_pop,predict_pop)*length(real_pop))/sum(real_pop)

#mae
mae(real_pop,predict_pop)

# %mae
(mae(real_pop,predict_pop) * length(real_pop))/sum(real_pop)
```
#3 visualize predicted pop and the real one  
```{r}

plot(density(sbl_predicted_df_adj_loa$loa_true_pop),col="black")+
lines(density(sbl_predicted_df_adj_loa$pred_loa_pop),col="red")
```

# plot the statistic residual
```{r}
plot(density(sbl_predicted_df_adj_loa$loa_true_pop-sbl_predicted_df_adj_loa$pred_loa_pop),col="black")
```


