---
title: "2_moa2sbl_rf"
author: "zhendong yuan"
date: "2019/10/11"
output: html_document
---


```{r setup, results='hide', message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("h2o")
library("ggplot2")
library("rsample") 
library("Metrics")
library("BBmisc")
library("rgdal")
```

# load pop
```{r} 
# at here the scaled_moa_feature.Rdata stands for only scaled populaiton 
moa_data = get(load("/home/z/zhendong/london/enriched_feature/moa_enriched_covariates.Rdata"))

```

# sperate training and testing set
```{r}
#h2o.shutdown(prompt = TRUE)
h2o.no_progress()
h2o.init(max_mem_size = "20g")
# Create training (70%) and test (30%) 
# Use set.seed for reproducibility
set.seed(123)
moa_data.h2o = as.h2o(moa_data)
split <- h2o.splitFrame(moa_data.h2o,c(0.7),seed=123)
train.h2o <- h2o.assign(split[[1]], "train.hex")   
test.h2o <- h2o.assign(split[[2]], "test.hex")  
y <- "population"
x <- setdiff(names(train.h2o), y)
```
# RF in loa level
# default setting base line
```{r}
# training basic GBM model with defaults
h2o.fit1 <- h2o.randomForest(
  training_frame = train.h2o,        ## the H2O frame for training
  x = x,
  y = y,
  ntrees = 5000, 
  max_depth = 100,
  stopping_rounds = 10
)

# assess model results
h2o.fit1
```

# randomDiscrete search
no need to make hyperparameter estimation, it already done by 2_moa2loa_enriched_RF

# set the grid searched/ fine tuned hyperparameters to get more robust result

```{r}
# train best model
h2o.final <- h2o.randomForest(
  x = x,
  y = y,
  nfolds = 5,
  training_frame = train.h2o,
  mtries = -1,
  nbins = 30,
   ntrees = 7000, 
  max_depth = 150,
  sample_rate = 0.8,
  stopping_rounds = 20,
  seed = 123
  
)

# model stopped after xx trees
h2o.final@parameters$ntrees

# cross validated RMSE
h2o.rmse(h2o.final, xval = TRUE)
```

# varaibles important
```{r}
h2o.varimp_plot(h2o.final, num_of_features = 10)
```

# prediction in moa's test dataset.
```{r}
# evaluate performance on new data
h2o.performance(model = h2o.final, newdata = test.h2o)

test_moa_predict = as.data.frame(h2o.predict(h2o.final, test.h2o))


predict_pop = test_moa_predict$predict
real_pop =  as.data.frame(test.h2o)$population

# rmse
rmse(real_pop,predict_pop)

# %rmse
(rmse(real_pop,predict_pop)*length(real_pop))/sum(real_pop)

#mae
mae(real_pop,predict_pop)

# %mae
(mae(real_pop,predict_pop) * length(real_pop))/sum(real_pop)

```

# test directly apply on SB
```{r}
# read loa feature table

sb_data= get(load("/home/z/zhendong/london/enriched_feature/sbl_enriched_covariates.Rdata"))
summary(sb_data)
# convert test set to h2o object
sb_transferred.h2o <- as.h2o(sb_data)

sb_transferred_predict = as.data.frame(h2o.predict(h2o.final, sb_transferred.h2o))

```




#1 denormorize it and aggregate to loa
```{r}
loa_or_data= get(load("/home/z/zhendong/london/Rspace/loa_data.Rdata"))
sum_or_pop = sum(loa_or_data$population)

sum_sb_transferred_predict = sum(sb_transferred_predict$predict)
rough_ratio_red = sum_or_pop/sum_sb_transferred_predict
# ratio transferred
rough_ratio_red

descaled_sb_transferred_predict = sb_transferred_predict$predict*rough_ratio_red
sum(descaled_sb_transferred_predict)

# aggreagete into loa according to loa_id
sb_shape <- readOGR("/home/z/zhendong/london/enriched_feature/street_blocksl_withinloa.shp")
colnames(sb_shape@data)
class(sb_shape@data$LSOA11CD)
sb_shape@data$LSOA11CD = as.character(sb_shape@data$LSOA11CD)
descaled_sb_transferred_predict_loa = NULL
descaled_sb_transferred_predict_loa = as.data.frame(cbind(descaled_sb_transferred_predict,sb_shape@data$LSOA11CD))
colnames(descaled_sb_transferred_predict_loa) = c("predicted_pop","LSOA11CD")
descaled_sb_transferred_predict_loa$predicted_pop = as.numeric(as.character(descaled_sb_transferred_predict_loa$predicted_pop))
descaled_sb_transferred_predict_loa$LSOA11CD = as.character(descaled_sb_transferred_predict_loa$LSOA11CD)



predicted_agg_sb2loa = descaled_sb_transferred_predict_loa %>% 
  group_by(LSOA11CD) %>% 
  summarise(predicted_pop = sum(predicted_pop))


raw_pop_loa= get(load("/home/z/zhendong/london/Rspace/pop/raw_pop_london_low.Rdata"))

raw_pop_loa = raw_pop_loa[,c(1,3)]
raw_pop_loa$AreaCodes = as.character(raw_pop_loa$AreaCodes)

pop_london_low = merge(raw_pop_loa,predicted_agg_sb2loa, by.x = "AreaCodes", by.y = "LSOA11CD", all = TRUE)

summary(pop_london_low)

pop_london_low[is.na(pop_london_low)] <- 0


```
#1 estimate accuracy 
```{r}
rmse(pop_london_low$population,pop_london_low$predicted_pop)
(rmse(pop_london_low$population,pop_london_low$predicted_pop)*length(pop_london_low$population))/sum(pop_london_low$population)
mae(pop_london_low$population,pop_london_low$predicted_pop)
(mae(pop_london_low$population,pop_london_low$predicted_pop) * length(pop_london_low$population))/sum(pop_london_low$population)
```
#1 visualize predicted pop and the real one  
```{r}
plot(density(pop_london_low$population),col="black")
lines(density(pop_london_low$predicted_pop),col="red")
```


#1 check the predicted pop and the real one pearson   correlation
```{r}
plot(pop_london_low$population,pop_london_low$predicted_pop, xlim = c(0,9000),ylim = c(0,9000))
abline(0,1, col='blue')
summary(pop_london_low)
# pearson
cor_matrix_1 <- cor(pop_london_low[c("population","predicted_pop")], use = 'complete.obs')
cor_matrix_1


```

#1 check residual
```{r}
residul = pop_london_low$population - pop_london_low$predicted_pop
plot(density(residul))
abline(v = 0, col='blue')
plot(pop_london_low$population,residul)

```

#3. test different aggreagation methods 
apply the restriction of moa 
```{r}

sbl_predicted_df = descaled_sb_transferred_predict_loa
colnames(sbl_predicted_df) = c("pred_pop","LSOA11CD") 

london_low <- readOGR("/home/z/zhendong/london/LSOA_2011_London_3857.shp")

london_low_df = london_low@data[,c(1,3)]
sbl_predicted_df_lm = merge(sbl_predicted_df,london_low_df, by= "LSOA11CD", all.x = TRUE)

raw_pop_moa= get(load("/home/z/zhendong/london/Rspace/pop/raw_pop_mid_london.Rdata"))
raw_pop_moa = raw_pop_moa[,c(1,3)]

# aggreagte sb to moa
sbl_predicted_df_lm_agg = sbl_predicted_df_lm %>% 
  group_by(MSOA11CD) %>% 
  summarise(predicted_pop = sum(pred_pop))
sbl_predicted_df_lm_agg$MSOA11CD = as.character(sbl_predicted_df_lm_agg$MSOA11CD)
sbl_predicted_df_lm_agg = as.data.frame(sbl_predicted_df_lm_agg)
raw_pop_moa["MSOA11CD"] = raw_pop_moa$AreaCodes
 raw_pop_moa$AreaCodes = NULL
sbl_predicted_df_lm_agg_temp = merge(sbl_predicted_df_lm_agg,raw_pop_moa,all.x = TRUE)
sbl_predicted_df_lm_agg_temp["ad_ratio"] = sbl_predicted_df_lm_agg_temp$population/sbl_predicted_df_lm_agg_temp$predicted_pop
# adjust by pop in moa
sbl_predicted_df_adj = NULL
sbl_predicted_df_adj = merge(sbl_predicted_df_lm,sbl_predicted_df_lm_agg_temp, by= "MSOA11CD", all.x = TRUE)
sbl_predicted_df_adj$pred_pop = sbl_predicted_df_adj$pred_pop*sbl_predicted_df_adj$ad_ratio
colnames(sbl_predicted_df_adj) = c("MSOA11CD","LSOA11CD","pred_pop","sum_sbl_pred_pop","moa_pop","adj_ratio")


# write the sb prediction into shape
#sb_shape@data["pred_adj_pop"] = sbl_predicted_df_adj$pred_pop
#writeOGR(sb_shape,dsn ="/home/z/zhendong/london/enriched_feature/moa2sbl",layer = "direactly_predicted_adj_sbl",driver = "ESRI Shapefile")


# aggreagete to loa
sbl_predicted_df_adj_loa = sbl_predicted_df_adj %>% 
  group_by(LSOA11CD) %>% 
  summarise(pred_loa_pop = sum(pred_pop))

# merge with true pop in loa 
raw_pop_loa= get(load("/home/z/zhendong/london/Rspace/pop/raw_pop_london_low.Rdata"))
raw_pop_loa["LSOA11CD"] = raw_pop_loa$AreaCodes
sbl_predicted_df_adj_loa = merge(sbl_predicted_df_adj_loa,raw_pop_loa, by= "LSOA11CD", all.x = TRUE)
colnames(sbl_predicted_df_adj_loa)
sbl_predicted_df_adj_loa = sbl_predicted_df_adj_loa[,c(1:5)]
colnames(sbl_predicted_df_adj_loa)[5] = "loa_true_pop"

summary(sbl_predicted_df_adj_loa)



#london_low_shp = merge(sbl_predicted_df_adj_loa,london_low@data,by = "LSOA11CD",all.y= TRUE)
#london_low@data = london_low_shp
#writeOGR(london_low,dsn ="/home/z/zhendong/london/enriched_feature/moa2sbl",layer = "direactly_sbl_predicted_adj_loa",driver = "ESRI Shapefile")

```
#3.  esitmate accuracy 
```{r}
predict_pop = sbl_predicted_df_adj_loa$pred_loa_pop
real_pop =  sbl_predicted_df_adj_loa$loa_true_pop

summary(predict_pop)
summary(real_pop)

# rmse
rmse(real_pop,predict_pop)

# %rmse
(rmse(real_pop,predict_pop)*length(real_pop))/sum(real_pop)

#mae
mae(real_pop,predict_pop)

# %mae
(mae(real_pop,predict_pop) * length(real_pop))/sum(real_pop)



```
#3 visualize predicted pop and the real one  
```{r}
plot(density(sbl_predicted_df_adj_loa$loa_true_pop),col="black")+
lines(density(sbl_predicted_df_adj_loa$pred_loa_pop),col="red")
```


#3. check the predicted pop and the real one    correlation
```{r}
plot(sbl_predicted_df_adj_loa$loa_true_pop,sbl_predicted_df_adj_loa$pred_loa_pop, xlim = c(0,9000),ylim = c(0,9000))
abline(0,1, col='blue')
# pearson
cor_matrix_2 <- cor(sbl_predicted_df_adj_loa[c("loa_true_pop","pred_loa_pop")], use = 'complete.obs')
cor_matrix_2

```

#3. check residual
```{r}
residul = sbl_predicted_df_adj_loa$loa_true_pop - sbl_predicted_df_adj_loa$pred_loa_pop

plot(density(residul))
abline(v = 0, col='blue')

plot(sbl_predicted_df_adj_loa$loa_true_pop,residul)

```




