---
title: "multiclass_dasymetric"
author: "***********"
date: "2019/10/4"
output: html_document
---

```{r setup, results='hide', message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("rgdal")
library("sp")
library("dplyr")
library("Metrics")
```

## read data

```{r}
#popdensity_class <- readOGR("/home/z/***********/london/multiclassdsymetric/classdensity.shp")
popdensity_class <- readOGR("/data/***********Backup/london/multiclassdsymetric/classdensity.shp")
raw_pop_mid_london = get(load("/home/z/***********/london/Rspace/pop/raw_pop_mid_london.Rdata"))
moa_class_df = merge(popdensity_class@data,raw_pop_mid_london, by.x = "MSOA11CD" ,by.y = "AreaCodes")
moa_class_df["pop_densi_class"] = (moa_class_df$population*moa_class_df$AREA_RATIO)/moa_class_df$AREA_CLASS
```

# population density fraction of urbanization class u in county c
```{r}
sum_class_pop_densi = sum(moa_class_df$pop_densi_class)
moa_class_df["pop_d"] =(moa_class_df$pop_densi_class / sum_class_pop_densi)

moa_class_df[,c(1,5:12)] = NULL
# total fraction
moa_class <- readOGR("/data/***********Backup/london/multiclassdsymetric/lulc_london_temp.shp")
total_fraction_class = merge(moa_class@data, moa_class_df, by.x = "CLASS_NAME", by.y="CLASS_NAME", all.x=TRUE)


moa_f_sum = total_fraction_class %>%
  group_by(MSOA11CD) %>%
  summarize(F_sum = sum(pop_d*pop_densi_class))

total_fraction_class_temp = merge(total_fraction_class, moa_f_sum, by.x = "MSOA11CD", by.y="MSOA11CD", all.x=TRUE)

total_fraction_class["total_fraction_class"] = (total_fraction_class_temp$pop_d*total_fraction_class_temp$AREA_RATIO.x)/total_fraction_class_temp$F_sum

# final population
final = merge(total_fraction_class,raw_pop_mid_london,by.x = "MSOA11CD" ,by.y = "AreaCodes", all.x = TRUE)
total_fraction_class["final_pop"] = (final$total_fraction_class*final$population)

summary(total_fraction_class["final_pop"])
```
```{r}
total_fraction_class["final_pop_dens"] = (total_fraction_class$final_pop/total_fraction_class$AREA_CLASS.x)
summary(total_fraction_class$AREA_CLASS.x)
summary(total_fraction_class$final_pop_dens)
```
# plot the prediction on sbl
```{r}
plot(density(total_fraction_class$final_pop))
```

# output the prediction on street block level
```{r}
moa_class@data$LULC_OBJ_G = as.numeric(as.character(moa_class@data$LULC_OBJ_G))
total_fraction_class$LULC_OBJ_G = as.numeric(as.character(total_fraction_class$LULC_OBJ_G))
result_sbl = merge(moa_class,total_fraction_class, by = "LULC_OBJ_G")
save(result_sbl,file = "/data/***********Backup/london/multiclassdsymetric/result_sbl.Rdata")
```


## aggragated to loa for valdation
```{r}

loa_pred_pop = total_fraction_class %>%
  group_by(LOA_GID) %>%
  summarize(LOA_pred_pop = sum(final_pop))
summary(loa_pred_pop)

raw_pop_loa_london = get(load("/home/z/***********/london/Rspace/pop/raw_pop_london_low.Rdata"))
loa_pred_pop = merge(loa_pred_pop,raw_pop_loa_london[,c("AreaCodes","population")],all.y=TRUE,by.x = "LOA_GID",by.y="AreaCodes")
loa_pred_pop$LOA_pred_pop[is.na(loa_pred_pop$LOA_pred_pop)] <- 0
summary(loa_pred_pop)
```
## accuracy metrics
```{r}
rmse(loa_pred_pop$population,loa_pred_pop$LOA_pred_pop)
(rmse(loa_pred_pop$population,loa_pred_pop$LOA_pred_pop)*length(loa_pred_pop$population))/sum(loa_pred_pop$population)
mae(loa_pred_pop$population,loa_pred_pop$LOA_pred_pop)
(mae(loa_pred_pop$population,loa_pred_pop$LOA_pred_pop) * length(loa_pred_pop$population))/sum(loa_pred_pop$population)
```
# plot the density
```{r}
plot(density(loa_pred_pop$population),col="black")+
lines(density(loa_pred_pop$LOA_pred_pop),col="red")
```

## st_intersection with loa.
```{r}
lulc_london_loa <- readOGR("/home/z/***********/london/multiclassdsymetric/lulc_london_loa.shp")
# here
lulc_london_loa_final = merge(lulc_london_loa@data,total_fraction_class,by="LULC_OBJ_G",all.x = TRUE)

lulc_london_loa_final["pop_loa_obj"] =  lulc_london_loa_final$final_pop*lulc_london_loa_final$AREA_RATIO

loa_pred_pop = lulc_london_loa_final %>%
  group_by(LOA_GID.x) %>%
  summarize(LOA_pred_pop_2 = sum(pop_loa_obj))
summary(loa_pred_pop)

loa_pred_pop = merge(loa_pred_pop,raw_pop_loa_london[,c("AreaCodes","population")],all.y=TRUE,by.x = "LOA_GID.x",by.y="AreaCodes")
loa_pred_pop$LOA_pred_pop_2[is.na(loa_pred_pop$LOA_pred_pop_2)] <- 0
summary(loa_pred_pop)

```
# output

```{r}
london_low <- readOGR("/home/z/***********/london/LSOA_2011_London_3857.shp")
colnames(london_low@data)
london_low_output = merge(london_low,loa_pred_pop,by.x="LSOA11CD", by.y = "LOA_GID.x",all.x = TRUE)
colnames(london_low_output@data)
london_low_output@data[,c(2:14)] = NULL
colnames(london_low_output@data)
save(london_low_output, file="/home/z/***********/london/multiclassdsymetric/london_low_output.Rdata")
```

## accuracy metrics
```{r}
rmse(loa_pred_pop$population,loa_pred_pop$LOA_pred_pop_2)
(rmse(loa_pred_pop$population,loa_pred_pop$LOA_pred_pop_2)*length(loa_pred_pop$population))/sum(loa_pred_pop$population)
mae(loa_pred_pop$population,loa_pred_pop$LOA_pred_pop_2)
(mae(loa_pred_pop$population,loa_pred_pop$LOA_pred_pop_2) * length(loa_pred_pop$population))/sum(loa_pred_pop$population)
```
# visualize predicted pop and the real one  
```{r}

plot(density(loa_pred_pop$population),col="black")+
lines(density(loa_pred_pop$LOA_pred_pop_2),col="red")
```
# check the residual 
```{r}
plot(density(loa_pred_pop$population-loa_pred_pop$LOA_pred_pop_2),col="black")
```

#check the predicted pop and the real one    correlation
```{r}
plot(loa_pred_pop$population,loa_pred_pop$LOA_pred_pop_2, xlim = c(0,9000),ylim = c(0,9000))
abline(0,1, col='blue')
# pearson
cor_matrix_2 <- cor(loa_pred_pop[c("population","LOA_pred_pop_2")], use = 'complete.obs')
cor_matrix_2

```

#3. check residual
```{r}
residul = loa_pred_pop$population - loa_pred_pop$LOA_pred_pop_2

plot(density(residul))
abline(v = 0, col='blue')

plot(loa_pred_pop$population,residul)

```








