---
title: "moa2loa_enriched_adaptation_KLIEP"
author: "**********"
date: "2019/9/22"
output: html_document
---
```{r setup, results='hide', message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("h2o")
library("rsample") 
library("Metrics")
library("densratio")
library(corrplot)
```

# in this file, domain adaption will be test from moa2loa
# load source and target data

```{r}
# load loa and moa
target_data = get(load("/home/z/**********/london/enriched_feature/loa_enriched_covariates.Rdata"))
plot(density(target_data$B_NUMBER))
summary(target_data)
source_data = get(load("/home/z/**********/london/enriched_feature/moa_enriched_covariates.Rdata"))
plot(density(source_data$B_NUMBER))
summary(source_data)
```

# directly ratio esitimation 
```{r}
target_data$population = NULL
target_data_dropPop = target_data
target_data = NULL
source_data$population = NULL
source_data_dropPop = source_data
source_data = NULL
target_data_dropPop[is.na(target_data_dropPop)] <- 0
source_data_dropPop[is.na(source_data_dropPop)] <- 0
```

#
```{r}
#densratio_obj_uLSIF <- densratio(source_data_dropPop,target_data_dropPop,method = "KLIEP")

#density_ratio = densratio_obj_uLSIF$compute_density_ratio(source_data_dropPop)
weights = get(load("/home/z/**********/london/enriched_feature/moa2loa_weights_RuLSIF.Rdata"))
density_ratio = weights
summary(density_ratio)
plot(density(density_ratio))
```


# weighted model which trained and fine-tuned from moa
```{r}
# load loa and moa
source_data = get(load("/home/z/**********/london/enriched_feature/moa_enriched_covariates.Rdata"))
source_data["weights"] = density_ratio


# save this weights for further analysis
#save(source_data,file = "/home/z/**********/london/enriched_feature/moa2loa_enriched_weights_KLIEP.Rdata")

set.seed(123)
source_data_split <- initial_split(source_data, prop = .7)
source_data_train <- training(source_data_split)
source_data_test  <- testing(source_data_split)
source_data = NULL
h2o.no_progress()
h2o.init(max_mem_size = "10g")
source_moa_train.h2o <- as.h2o(source_data_train)
y <- "population"
x <- setdiff(names(source_moa_train.h2o), y)
# train final model
h2o.moa.final <- h2o.randomForest(
  
  x = x,
  y = y,
  
  training_frame = source_moa_train.h2o,
  weights_column = "weights",



 nfolds = 5,
  mtries = -1,
  nbins = 30,
   ntrees = 7000, 
  max_depth = 150,
  sample_rate = 0.8,
  stopping_rounds = 20



)

# model stopped after xx trees
h2o.moa.final@parameters$ntrees

# cross validated RMSE
h2o.rmse(h2o.moa.final, xval = TRUE)
```
# weighted model predict on moa level. validat in testing dataset
```{r}
source_data_test$weights = NULL
source_data_test.h2o <- as.h2o(source_data_test)
h2o.performance(model = h2o.moa.final, newdata = source_data_test.h2o)
source_data_test_population = source_data_test$population
source_data_test$population = NULL
source_data_test.h2o <- as.h2o(source_data_test)
# evaluate performance on loa
source_data_test_predict = as.data.frame(h2o.predict(h2o.moa.final, source_data_test.h2o))
source_data_test["population"] = source_data_test_population
```
```{r}
predict_pop = source_data_test_predict$predict
real_pop =  source_data_test$population

# rmse
rmse(real_pop,predict_pop)

# %rmse
(rmse(real_pop,predict_pop)*length(real_pop))/sum(real_pop)

#mae
mae(real_pop,predict_pop)

# %mae
(mae(real_pop,predict_pop) * length(real_pop))/sum(real_pop)
```
# weighted model applied to data at loa level
```{r}
target_data = get(load("/home/z/**********/london/enriched_feature/loa_enriched_covariates.Rdata"))

#target_data$population = NULL
target.loa.h2o <- as.h2o(target_data)
# evaluate performance on loa
weighted_loa_performance = h2o.performance(model = h2o.moa.final, newdata = target.loa.h2o)
weighted_loa_performance

(weighted_loa_performance@metrics$RMSE*length(target_data$population))/sum(target_data$population)
(weighted_loa_performance@metrics$mae*length(target_data$population))/sum(target_data$population)

```
# check the distribution of response
```{r}
target_data$population = NULL
target.loa.h2o <- as.h2o(target_data)
loa_transferred_predict = as.data.frame(h2o.predict(h2o.moa.final, target.loa.h2o))

target_data = get(load("/home/z/**********/london/enriched_feature/loa_enriched_covariates.Rdata"))
plot(density(target_data$population),col="black")+
  lines(density(source_data_train$population),col="green")+
lines(density(loa_transferred_predict$predict),col="red")+
lines(density(source_data_test_predict$predict),col= "blue")+
lines(density(source_data_test$population),col = "purple")

```
# check the residual correlation
```{r}
#target_data$population,loa_transferred_predict$predict

result_df = as.data.frame(cbind(loa_transferred_predict$predict,target_data$population))
colnames(result_df) = c("loa_transferred_predict","target_pop")
result_df.lm = lm(target_pop ~ loa_transferred_predict, data=result_df) 
plot(target_data$population,loa_transferred_predict$predict,xlim = c(0,1),ylim = c(0,1))
abline(0,1, col='blue')
abline(result_df.lm,col="red")
summary(result_df.lm)
```



# denormalization
```{r}
moa_or_data= get(load("/home/z/**********/london/Rspace/pop/raw_pop_mid_london.Rdata"))
loa_or_data= get(load("/home/z/**********/london/Rspace/pop/raw_pop_london_low.Rdata"))
sum_or_pop = sum(moa_or_data$population)

sum_loa_transferred_predict = sum(loa_transferred_predict$predict)
rough_ratio_red = sum_or_pop/sum_loa_transferred_predict
descaled_loa_transferred_predict = loa_transferred_predict$predict*rough_ratio_red
# ratio transferred
rough_ratio_red
# estimate rmse
rmse(loa_or_data$population,descaled_loa_transferred_predict)
(rmse(loa_or_data$population,descaled_loa_transferred_predict)*length(loa_or_data$population))/sum(loa_or_data$population)
mae(loa_or_data$population,descaled_loa_transferred_predict)
(mae(loa_or_data$population,descaled_loa_transferred_predict) * length(loa_or_data$population))/sum(loa_or_data$population)


```

#plot
```{r}
plot(density(loa_or_data$population),col="black")
lines(density(moa_or_data$population),col="green")
lines(density(descaled_loa_transferred_predict),col="red")
#lines(density(descaled_source_data_test_predict),col = "blue")
```


# check the predict vs true 
```{r}
#target_data$population,loa_transferred_predict$predict

de_nor_result_df = as.data.frame(cbind(descaled_loa_transferred_predict,loa_or_data$population))
colnames(de_nor_result_df) = c("descaled_loa_transferred_predict","loa_pop")
de_result_df.lm = lm(loa_pop ~ descaled_loa_transferred_predict, data=de_nor_result_df) 
plot(loa_or_data$population,descaled_loa_transferred_predict,xlim = c(0,10000),ylim = c(0,10000))
abline(0,1, col='blue')
abline(de_result_df.lm,col="red")
summary(de_result_df.lm)
```
```{r}
cor_matrix_2 <- cor(de_nor_result_df, use = 'complete.obs')
cor_matrix_2
```

# check the correlation with B_area_sum
```{r}
residuls_tf = abs(descaled_loa_transferred_predict-loa_or_data$population)

plot(density(residuls_tf))
loa_data = get(load("/home/z/**********/london/enriched_feature/loa_enriched_covariates.Rdata"))
plot(loa_data$B_AREA_SUM,residuls_tf)

```


```{r}

residul_df = as.data.frame(cbind(loa_data,descaled_loa_transferred_predict,loa_or_data$population,residuls_tf))
colnames(residul_df)[c(37:39)] = c("pred_pop","true_pop","resi_pop")
cor_matrix <- cor(residul_df, use = 'complete.obs')
corrplot.mixed(cor_matrix, lower = "circle", upper = "number", tl.pos = "lt", diag = "u")
```

