---
title: "enriched_moa2sbl_KLIEP"
author: "********"
date: "2019/10/12"
output: html_document
---

```{r setup, results='hide', message=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("dplyr")
library("h2o")
library("rsample") 
library("Metrics")
library("densratio")
library("rgdal")
```

# in this file, domain adaption will be test from moa2loa
# load source and target data

```{r}
# load loa and moa
target_data = get(load("/home/z/********/london/enriched_feature/sbl_enriched_covariates.Rdata"))
plot(density(target_data$B_NUMBER))
colnames(target_data)
source_data = get(load("/home/z/********/london/enriched_feature/moa_enriched_covariates.Rdata"))
plot(density(source_data$B_NUMBER))
colnames(source_data)


```

# directly ratio esitimation 
```{r}
target_data$population = NULL
target_data_dropPop = target_data
target_data = NULL
source_data$population = NULL
source_data_dropPop = source_data
source_data = NULL
target_data_dropPop[is.na(target_data_dropPop)] <- 0
source_data_dropPop[is.na(source_data_dropPop)] <- 0
```

```{r}
summary(source_data_dropPop)
summary(target_data_dropPop)
```

#
```{r}
densratio_obj_uLSIF <- densratio(source_data_dropPop,target_data_dropPop,method = "KLIEP")

density_ratio = densratio_obj_uLSIF$compute_density_ratio(source_data_dropPop)
summary(density_ratio)
plot(density(density_ratio))
```


# weighted model which trained and fine-tuned from moa
```{r}
# load loa and moa
source_data = get(load("/home/z/********/london/enriched_feature/moa_enriched_covariates.Rdata"))
source_data["weights"] = density_ratio


# save this weights for further analysis
save(source_data,file = "/home/z/********/london/enriched_feature/moa2sbl_enriched_weights_KLIEP.Rdata")

set.seed(123)
source_data_split <- initial_split(source_data, prop = .7)
source_data_train <- training(source_data_split)
source_data_test  <- testing(source_data_split)
source_data = NULL
h2o.no_progress()
h2o.init(max_mem_size = "10g")
source_moa_train.h2o <- as.h2o(source_data_train)
y <- "population"
x <- setdiff(names(source_moa_train.h2o), y)
# train final model
h2o.moa.final <- h2o.randomForest(
  
  x = x,
  y = y,
  
  training_frame = source_moa_train.h2o,
  weights_column = "weights",



 nfolds = 5,
  mtries = -1,
  nbins = 30,
   ntrees = 7000, 
  max_depth = 150,
  sample_rate = 0.8,
  stopping_rounds = 20



)

# model stopped after xx trees
h2o.moa.final@parameters$ntrees

# cross validated RMSE
h2o.rmse(h2o.moa.final, xval = TRUE)

# plot model.
h2o.moa.final
```
# weighted model predict on moa level. validat in testing dataset
```{r}
source_data_test$weights = NULL
source_data_test.h2o <- as.h2o(source_data_test)
h2o.performance(model = h2o.moa.final, newdata = source_data_test.h2o)
source_data_test_population = source_data_test$population
source_data_test$population = NULL
source_data_test.h2o <- as.h2o(source_data_test)
# evaluate performance on loa
source_data_test_predict = as.data.frame(h2o.predict(h2o.moa.final, source_data_test.h2o))
source_data_test["population"] = source_data_test_population
```
```{r}
predict_pop = source_data_test_predict$predict
real_pop =  source_data_test$population

# rmse
rmse(real_pop,predict_pop)

# %rmse
(rmse(real_pop,predict_pop)*length(real_pop))/sum(real_pop)

#mae
mae(real_pop,predict_pop)

# %mae
(mae(real_pop,predict_pop) * length(real_pop))/sum(real_pop)
```

# weighted model applied to data at sbl level
```{r}
target_data = get(load("/home/z/********/london/enriched_feature/sbl_enriched_covariates.Rdata"))
colnames(target_data)
#target_data$population = NULL
target.loa.h2o <- as.h2o(target_data)
sb_transferred_predict = as.data.frame(h2o.predict(h2o.moa.final, target.loa.h2o))
plot(density(sb_transferred_predict$predict))

```


#1 denormalize it 
```{r}
loa_or_data= get(load("/home/z/********/london/Rspace/loa_data.Rdata"))
moa_or_data= get(load("/home/z/********/london/Rspace/moa_data.Rdata"))
sum_or_pop = sum(moa_or_data$population)
print("sum true pop")
sum_or_pop
sum_sb_transferred_predict = sum(sb_transferred_predict$predict)
rough_ratio_red = sum_or_pop/sum_sb_transferred_predict
# ratio transferred
print("ratio transferred")
rough_ratio_red

descaled_sb_transferred_predict = sb_transferred_predict$predict*rough_ratio_red
print("sum up descaled sb predicted pop")
sum(descaled_sb_transferred_predict)

# aggreagete into loa according to loa_id
sb_shape <- readOGR("/home/z/********/london/enriched_feature/street_blocksl_withinloa.shp")
colnames(sb_shape@data)
class(sb_shape@data$LSOA11CD)
sb_shape@data$LSOA11CD = as.character(sb_shape@data$LSOA11CD)
descaled_sb_transferred_predict_loa = NULL
descaled_sb_transferred_predict_loa = as.data.frame(cbind(descaled_sb_transferred_predict,sb_shape@data$LSOA11CD))
colnames(descaled_sb_transferred_predict_loa) = c("predicted_pop","LSOA11CD")
descaled_sb_transferred_predict_loa$predicted_pop = as.numeric(as.character(descaled_sb_transferred_predict_loa$predicted_pop))
descaled_sb_transferred_predict_loa$LSOA11CD = as.character(descaled_sb_transferred_predict_loa$LSOA11CD)

# store this 
#sb_shape@data["pred_pop"] = descaled_sb_transferred_predict
#temp = sb_shape@data
#save(temp,file="/home/z/********/london/enriched_feature/moa2sbl/enriched_sbl_predict_pop_KLIEP_1.Rdata")

predicted_agg_sb2loa = descaled_sb_transferred_predict_loa %>% 
  group_by(LSOA11CD) %>% 
  summarise(predicted_pop = sum(predicted_pop))

raw_pop_loa= get(load("/home/z/********/london/Rspace/pop/raw_pop_london_low.Rdata"))

raw_pop_loa = raw_pop_loa[,c(1,3)]
raw_pop_loa$AreaCodes = as.character(raw_pop_loa$AreaCodes)

pop_london_low = merge(raw_pop_loa,predicted_agg_sb2loa, by.x = "AreaCodes", by.y = "LSOA11CD", all = TRUE)

summary(pop_london_low)

pop_london_low[is.na(pop_london_low)] <- 0
print("sum up descaled loa predicted pop")
sum(pop_london_low$predicted_pop)
print("sum up true loa pop")
sum(pop_london_low$population)
```
#1
```{r}
rmse(pop_london_low$population,pop_london_low$predicted_pop)
(rmse(pop_london_low$population,pop_london_low$predicted_pop)*length(pop_london_low$population))/sum(pop_london_low$population)
mae(pop_london_low$population,pop_london_low$predicted_pop)
(mae(pop_london_low$population,pop_london_low$predicted_pop) * length(pop_london_low$population))/sum(pop_london_low$population)
```


#1 visualize predicted pop and the real one  
```{r}

plot(density(pop_london_low$population),col="black")+
  lines(density(source_data_train$population),col="green")+
lines(density(pop_london_low$predicted_pop),col="red")+
lines(density(source_data_test_predict$predict),col= "blue")+
lines(density(source_data_test$population),col = "purple")
```

#3. test different aggreagation methods 
apply the restriction of moa 

```{r}
sb_transferred_predict = get(load("/home/z/********/london/enriched_feature/moa2sbl/predicted_result_temp.Rdata"))
sb_shape <- readOGR("/home/z/********/london/enriched_feature/street_blocksl_withinloa.shp")
sb_shape@data$sb_id = c(1:length(sb_shape@data$AREA))
sbl_predicted_data = data.frame(pred_sbl_pop = sb_transferred_predict$predict,LSOA11CD=sb_shape@data$LSOA11CD, sb_id=sb_shape@data$sb_id)


london_low <- readOGR("/home/z/********/london/LSOA_2011_London_3857.shp")
london_low_df = london_low@data[,c(1,3)]

sbl_predicted_df_lm = merge(sbl_predicted_data,london_low_df, by= "LSOA11CD", all.x = TRUE)



# aggreagte sb to moa
sbl_predicted_df_lm_agg = sbl_predicted_df_lm %>%
  group_by(MSOA11CD) %>% 
  summarise(predicted_pop_sum_moa = sum(pred_sbl_pop))

raw_pop_moa= get(load("/home/z/********/london/Rspace/pop/raw_pop_mid_london.Rdata"))
raw_pop_moa = raw_pop_moa[,c(1,3)]
colnames(raw_pop_moa) = c("MSOA11CD","moa_true_population")
summary(raw_pop_moa)

sbl_predicted_df_lm_agg_temp = merge(sbl_predicted_df_lm_agg,raw_pop_moa,by = "MSOA11CD",all.x = TRUE)
sbl_predicted_df_lm_agg_temp["ad_ratio"] = sbl_predicted_df_lm_agg_temp$moa_true_population/sbl_predicted_df_lm_agg_temp$predicted_pop_sum_moa
# adjust by pop in moa
sbl_predicted_df_adj = NULL
sbl_predicted_df_adj = merge(sbl_predicted_df_lm,sbl_predicted_df_lm_agg_temp, by= "MSOA11CD", all.x = TRUE)

sbl_predicted_df_adj$pred_adj_pop = sbl_predicted_df_adj$pred_sbl_pop * sbl_predicted_df_adj$ad_ratio

temp_final = merge(sb_shape@data,sbl_predicted_df_adj[,c("pred_sbl_pop","pred_adj_pop","sb_id")],by="sb_id",all.x=TRUE)

#sb_shape@data=temp_final
#writeOGR(sb_shape,dsn ="/home/z/********/london/enriched_feature/moa2sbl",layer = "predicted_adj_sbl",driver = "ESRI Shapefile")
```

# aggreagete to loa
```{r}
# aggreagete to loa
sbl_predicted_df_adj_loa = sbl_predicted_df_adj %>% 
  group_by(LSOA11CD) %>% 
  summarise(pred_loa_pop = sum(pred_adj_pop))

# merge with true pop in loa 
raw_pop_loa= get(load("/home/z/********/london/Rspace/pop/raw_pop_london_low.Rdata"))
raw_pop_loa["LSOA11CD"] = raw_pop_loa$AreaCodes
sbl_predicted_df_adj_loa = merge(sbl_predicted_df_adj_loa,raw_pop_loa, by= "LSOA11CD", all.x = TRUE)
colnames(sbl_predicted_df_adj_loa)
sbl_predicted_df_adj_loa = sbl_predicted_df_adj_loa[,c(1:5)]


colnames(sbl_predicted_df_adj_loa)[5] = "loa_true_pop"
sbl_predicted_df_adj_loa["residual"] = sbl_predicted_df_adj_loa$loa_true_pop-sbl_predicted_df_adj_loa$pred_loa_pop
summary(sbl_predicted_df_adj_loa)

london_low_shp = merge(sbl_predicted_df_adj_loa,london_low@data,by = "LSOA11CD",all.y= TRUE)
#london_low@data = london_low_shp
#writeOGR(london_low,dsn ="/home/z/********/london/enriched_feature/moa2sbl",layer = "sbl_predicted_adj_loa",driver = "ESRI Shapefile")
```

#3.  esitmate accuracy 
```{r}
predict_pop = sbl_predicted_df_adj_loa$pred_loa_pop
real_pop =  sbl_predicted_df_adj_loa$loa_true_pop

# rmse
rmse(real_pop,predict_pop)

# %rmse
(rmse(real_pop,predict_pop)*length(real_pop))/sum(real_pop)

#mae
mae(real_pop,predict_pop)

# %mae
(mae(real_pop,predict_pop) * length(real_pop))/sum(real_pop)
```
#3 visualize predicted pop and the real one  
```{r}

plot(density(sbl_predicted_df_adj_loa$loa_true_pop),col="black")+
lines(density(sbl_predicted_df_adj_loa$pred_loa_pop),col="red")
```


```{r}
plot(hist(sbl_predicted_df_adj_loa$loa_true_pop),col="black")
#+lines(density(sbl_predicted_df_adj_loa$pred_loa_pop),col="red")
```